
---
title: "Data 607 Week 3 - Tuberculosis Rates"
author: "Armenoush Aslanian-Persico"
output: html_document
css: C:/data/style.css
---

### 0. Optional: <a href="https://raw.githubusercontent.com/spsstudent15/msda20161/master/style.css">Download CSS Stylesheet here.</a>
#### Save into working directory.
```{r}
getwd()
```

### 1. Load packages
```{r warning=FALSE, echo=FALSE}
library('rvest')
library('devtools')
library('RCurl')
library('plyr')
library('dplyr')
library('knitr')
```

### 2. Create a CSV file using SQL.

#### Resulting CSV file was uploaded to GitHub for Step 2.
```sql
`r paste(readLines('C:/data/tb.sql'), collapse = '\n')`
```

### 3. Import CSV file from GitHub.
#### Indicate "header=F" for missing header row.
```{r}
x <- getURL("https://raw.githubusercontent.com/spsstudent15/msda20161/master/tb3.csv")
dftb<-data.frame(read.csv(text=x, header=F, stringsAsFactors = F))

```


### 4. View data frame dimensions, head, structure.
```{r}
dim(dftb)
head(dftb)
str(dftb)
```
#### This returns 1900 records for 9 variables and shows we need to convert character columns to numeric.


### 5. Convert character columns to numeric. This converts NA values to 0.
```{r}
dftb$V4<-as.numeric(dftb$V4)
dftb$V5<-as.numeric(dftb$V5)
dftb$V6<-as.numeric(dftb$V6)
dftb$V7<-as.numeric(dftb$V7)
dftb$V9<-as.numeric(dftb$V9)
```
### 6. View head and structure.
```{r}
str(dftb)
head(dftb)
```


### 7. Group by year, create bar plot.
#### Check against Excel and SQL:

#### 1995: 494742 total cases
#### 1996: 615510 total cases
```{r}
grpyear<-aggregate(cbind(V7) ~ V3, data = dftb, sum)
grpyear

barplot(
  as.numeric(grpyear$V7/1000000), 
  names=grpyear$V3, 
  xlab="Year",
  ylab="Cases in Millions",
  main="Worldwide Tuberculosis Cases by Year", 
  border="black", 
  col="blue", 
  ylim=c(0,5))

```


### 8. Rename column headers.
```{r}
colnames(dftb) <-c(
  "CID",
  "COUNTRY",
  "YEAR",
  "CHILD",
  "ADULT",
  "ELDERLY",
  "TOTAL", 
  "COUNTRY_POPULATION", 
  "RATE")

```

### 9. What kind of analysis and reporting could be done with this data?
#### We could see trends of increasing rates in specific countries and adjust public policy and funding to address those issues. We could also see which countries are missing data.

### 10. Table output.
#### Format table with Kable package.
```{r}
tb3s <- subset(dftb,select= c( 2,3,9))
summary(tb3s)
kable(tb3s,align="c",caption="Tuberculosis Infection Rates Per Capita by Country and Year")
```

